```{r col_ctry}
ctry <- "COL"
ctry2 <- "Colombia"
```

# Colombia

```{r col_pkgs}
library(oecdcountryprofiles)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(oecdplot)
library(knitr)
library(readxl)
library(janitor)
library(stringr)
library(readr)
library(tintin)
library(scales)

load_oecd_fonts()

missing_reg <- c("ITG2F")
excl_countries <- c("LUX", "IRL")

clrs <- tintin_colours$the_black_island
clrs2 <- tintin_colours$the_blue_lotus
clrs3 <- tintin_colours$the_red_sea_sharks
clrs4 <- tintin_colours$the_calculus_affair
clrs5 <- tintin_colours$red_rackhams_treasure
```

<table style="margin-left:66.1pt;border-collapse:collapse" cellspacing="0" cellpadding="0" border="0">
<tbody>
<tr>
<td style="width:467.5pt;border:solid windowtext 1.0pt;padding:0cm 5.4pt 0cm 5.4pt" width="623" valign="top">
<p class="MsoNormal" style="background:#fff9e6">
<b><span style="font-size:9.0pt;color:#212529">OECD Regional Outlook</span></b><u></u><u></u></p>
<p class="MsoNormal" style="margin:1pt;background:white"><span style="font-size:9.0pt;color:#212529">The OECD
<a href="https://www.oecd-ilibrary.org/urban-rural-and-regional-development/oecd-regional-outlook_2dafc8cf-en" target="_blank"><i><span style="color:#0d6efd">Regional Outlook</span></i></a> reviews recent trends, policy developments, and prospects across OECD regions, including the underlying causes driving regional inequalities in performance and well-being. The report offers evidence,
 guidance and policy recommendations on how to improve competitiveness and productivity, promote inclusive growth, accelerate the net-zero transition and raise well-being standards through effective regional development policy and multi-level governance.</span><u></u><u></u></p>
</td>
</tr>
</tbody>
</table>


## Overview

```{r col_tbl}
read_html_table(ctry)
```

## Regional inequality trends

```{r col_fig1}
iso3_g20 <- read_csv("data/ISO_3_G20.csv") %>%
  clean_names() %>%
  rename(iso3 = iso_3)

top_bottom_c <- read_excel("data/top_bottom_gdppc_w_agg.xlsx") %>%
  clean_names()

gdp_pop_00_20 <- read_csv("data/gdp_pop_00_20_imputed_agg_l.csv") %>%
  clean_names() %>%
  select(-x1) %>%
  left_join(iso3_g20, by = "iso3") %>%
  filter(!reg_id %in% missing_reg, !iso3 %in% excl_countries) %>%
  mutate(
    metro_large = ifelse(metro_large != "Metropolitan large", "Non-large", metro_large),
    metro_far = ifelse(metro_far != "Far from a FUA>250K", "Close", metro_far)
  ) %>%
  bind_rows(
    read_csv("data/gdp_pop_tl2_00_20_imputed_l.csv") %>%
      clean_names() %>%
      select(time, reg_id, iso3, gdp, pop, gdp_pc, regional_name_eng = reg_name)
  )

summary_theil_tl2 <- read_csv("data/theil_gdppc_country_00_20_wCHE.csv") %>%
  clean_names() %>%
  filter(level == "gdppc_tl2") %>%
  select(-c(x1, level))

summary_stats <- gdp_pop_00_20 %>%
  group_by(iso3, time) %>%
  summarise(mean_gdp_pc = mean(gdp_pc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(top_bottom_c, by = c("iso3", "time")) %>%
  mutate(
    top_bottom = top_20_gdppc / bottom_20_gdppc,
    top_mean = top_20_gdppc / mean_gdp_pc,
    bottom_mean = bottom_20_gdppc / mean_gdp_pc
  )

summary_stats <- summary_stats %>%
  full_join(summary_theil_tl2, by = c("iso3", "time"))

# summary_stats %>%
#   filter(iso3 == ctry)

# summary_theil_tl2 %>%
#   filter(iso3 == ctry)

index_2000 <- summary_stats %>%
  filter(time == 2000) %>%
  pivot_wider(
    id_cols = "iso3",
    names_from = "time",
    values_from = -c(time, iso3)
  )

summary_stats_index <- summary_stats %>%
  full_join(index_2000, by = "iso3") %>%
  mutate(
    index_gdppc = mean_gdp_pc / mean_gdp_pc_2000,
    index_bottom = bottom_mean / bottom_mean_2000,
    index_top = top_mean / top_mean_2000,
    index_theil = theil / theil_2000,
    index_top_mean = top_mean / top_mean_2000,
    index_bottom_mean = bottom_mean / bottom_mean_2000,
    index_top_bottom = top_bottom / top_bottom_2000
  ) %>%
  select(
    iso3,
    time,
    index_gdppc,
    index_bottom,
    index_top_mean,
    index_top_bottom,
    index_bottom_mean,
    index_top,
    index_theil
  )

summary_wide <- summary_stats_index %>%
  pivot_longer(cols = -c(time, iso3), names_to = "index") %>%
  mutate(
    labels_index = case_when(
      index == "index_bottom" ~ "Bottom 20%",
      index == "index_top" ~ "Top 20%",
      index == "index_bottom_mean" ~ "Bottom 20%/Mean",
      index == "index_top_mean" ~ "Top 20%/Mean",
      index == "index_theil" ~ "Theil index",
      index == "index_gdppc" ~ "Mean GDP per capita",
      index == "index_top_bottom" ~ "Top 20%/Bottom 20%"
    )
  ) %>%
  left_join(iso3_g20, by = "iso3")

lev_2 <- c(
  "Theil index",
  "Top 20%/Mean",
  "Bottom 20%/Mean",
  "Mean GDP per capita"
)

fig1 <- summary_wide %>%
  filter(iso3 == ctry, labels_index %in% lev_2) %>%
  pivot_wider(
    names_from = index,
    values_from = value
  ) %>%
  ggplot(aes(x = time)) +
  geom_line(aes(
    y = index_gdppc,
    colour = factor(labels_index, levels = lev_2)
  ), linetype = "dashed", linewidth = 1.2) +
  geom_line(aes(
    y = index_top_mean,
    colour = factor(labels_index, levels = lev_2)
  ), linetype = "dashed", linewidth = 1.2) +
  geom_line(aes(
    y = index_bottom_mean,
    colour = factor(labels_index, levels = lev_2)
  ), linetype = "dashed", linewidth = 1.2) +
  geom_line(aes(
    y = index_theil,
    colour = factor(labels_index, levels = lev_2)
  ), linewidth = 1.2) +
  scale_colour_manual(values = clrs2) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(
    title = "Figure 1: Trends in GDP per capita inequality indicators,\nTL2 OECD regions",
    x = "",
    y = "Statistic (2000=1)",
    linetype = "",
    colour = ""
  ) +
  theme_oecd(base_size = 10) +
  theme(plot.title = element_text(size = 13, hjust = 0, margin = margin(0, 0, 8, 0)))
```

```{r col_fig1_summary}
theil_chg <- summary_wide %>%
  filter(
    index == "index_theil",
    iso3 == ctry,
    time %in% range(time)
  ) %>%
  arrange(time) %>%
  mutate(diff = value - lag(value)) %>%
  filter(time == max(time)) %>%
  pull(diff)

theil_chg_txt <- ifelse(theil_chg > 0, "an increase", "a decline")

theil_max <- summary_wide %>%
  filter(
    index == "index_theil",
    iso3 == ctry
  ) %>%
  filter(value == max(value)) %>%
  pull(time) %>%
  as.integer()

polarization_pct <- summary_wide %>%
  filter(
    index == "index_top_mean",
    iso3 == ctry,
    time %in% range(time)
  ) %>%
  arrange(time) %>%
  mutate(diff = lead(value) - value) %>%
  filter(time == min(time)) %>%
  pull(diff) %>%
  round(3)

polarization_txt <- ifelse(polarization_pct > 0, "higher", "lower")

polarization_yrs <- summary_wide %>%
  filter(
    index == "index_top_mean",
    iso3 == ctry
  ) %>%
  filter(time %in% range(time)) %>%
  pull(time) %>%
  as.integer()

polarization_2_pct <- summary_wide %>%
  filter(
    index == "index_bottom_mean",
    iso3 == ctry,
    time %in% range(time)
  ) %>%
  arrange(time) %>%
  mutate(diff = lead(value) - value) %>%
  filter(time == min(time)) %>%
  pull(diff) %>%
  round(3)

polarization_2_txt <- ifelse(polarization_2_pct > 0, "higher", "lower")
```

`r ctry2` experienced `r theil_chg_txt` in the Theil index of GDP per capita over 2000-2020. Inequality reached its maximum in `r theil_max`.

<!-- Polarisation, as measured by the Top 20%/Mean ratio was `r paste(abs(polarization_pct), polarization_txt)` in `r min(polarization_yrs)` compared to `r max(polarization_yrs)`. Bottom divergence, as measured by the Bottom 20%/Mean ratio was `r paste(abs(polarization_2_pct), polarization_2_txt)` in the same period. -->

```{r col_fig1_1}
fig1
```

<div class="oecdnote">
**Note**: Based on 1 586 TL3 regions in 27 countries with available data (no TL3 data (continuous time series for more than 1 region) for Australia, Canada, Chile, Colombia, Costa Rica, Iceland, Ireland, Israel, Mexico, Luxembourg and Switzerland). Between Theil measures the dissimilarity of the national GDP per capita means with respect to the OECD average. Within Theil measures the dissimilarity between regional and national GDP per capita.<br>
**Source**: OECD Regional Database (2022).
</div>

<br>
<br>

```{r col_fig3}
# read ----

finp <- if (any(ctry %in% c("AUS", "CAN", "COL", "CHE", "CHL", "IRL", "MEX"))) {
  "data/countryprofile_option1_addon.xlsx"
} else {
  "data/countryprofile_option1.xlsx"
}

dp1 <- read_excel(finp, sheet = ctry) %>%
  clean_names()

# tidy ----

colnames(dp1) <- str_replace(colnames(dp1), tolower(ctry), "country")

dp11 <- dp1 %>%
  select(time, matches("country")) %>%
  pivot_longer(-time) %>%
  rename(value_country = value)

dp11 <- dp11 %>%
  mutate(
    name = case_when(
      name == "pw_lp_country" ~ "Lower half",
      name == "pw_hp_country" ~ "Upper half"
    )
  )

# plot ----

yrs <- sort(unique(dp11$time))
yrs <- seq(min(yrs), max(yrs), 2)

hpgrew <- dp11 %>%
  filter(
    name == "Upper half",
    time %in% c(min(yrs), max(yrs))
  ) %>%
  summarise(grew = 100 * (value_country - lag(value_country)) / lag(value_country)) %>%
  drop_na() %>%
  pull() %>%
  round(1)

lpgrew <- dp11 %>%
  filter(
    name == "Lower half",
    time %in% c(min(yrs), max(yrs))
  ) %>%
  summarise(grew = 100 * (value_country - lag(value_country)) / lag(value_country)) %>%
  drop_na() %>%
  pull() %>%
  round(1)

hpdiff <- hpgrew - lpgrew

hpmore <- ifelse(hpdiff > 0, "more", "less")

fig3_1 <- ggplot(dp11) +
  geom_line(aes(x = time, y = value_country, color = name), linewidth = 1.2) +
  theme_oecd(base_size = 10) +
  scale_colour_manual(values = clrs3[1:2]) +
  labs(
    x = "", y = "Productivity (2015 USD PPP)", colour = "Series",
    title = "Figure 2: Evolution of labour productivity, TL2 regions"
  ) +
  scale_x_continuous(labels = as.character(yrs), breaks = yrs) +
  scale_y_continuous(labels = scales::number_format()) +
  theme(plot.title = element_text(size = 13, hjust = 0, margin = margin(0, 0, 8, 0)))
```

In `r ctry2`, labour productivity in high-productivity regions grew by `r paste0(hpgrew, "%")` between `r min(yrs)` and `r max(yrs)`, `r paste0(abs(hpdiff))` percentage points `r hpmore` compared to low-productivity regions.

```{r col_fig3_2}
fig3_1
```

<div class="oecdnote">
**Note**: Labour productivity for high productivity (low productivity) regions is equal to the sum of Gross Value Added across high productivity (low productivity) regions in a country/year divided by the sum of Employment across high productivity (low productivity) regions in a country/year. Gross Value Added is expressed in USD 2015 PPP. Regions are at the TL2 level for Australia, Canada, Chile, Colombia, Ireland, Mexico, Norway, Switzerland, Turkey and the United States. A region is defined as high productivity if its productivity was equal or above the country median for at least two years out of the first four years of each region’ observation period, and low productivity otherwise. The starting year is the first available year for each country. <br>
**Source**: OECD Regional Database (2022).
</div>

<br>
<br>

## Recent policy developments

```{r col_txt}
read_html_text(ctry)
```

<br>
<br>

<table style="margin-left:66.1pt;border-collapse:collapse" cellspacing="0" cellpadding="0" border="0">
<tbody>
<tr>
<td style="width:467.5pt;border:solid windowtext 1.0pt;padding:0cm 5.4pt 0cm 5.4pt" width="623" valign="top">
<p class="MsoNormal" style="background:#d4edff"><b><span style="font-size:9.0pt;color:#212529">Territorial definitions</span></b><u></u><u></u></p>
<p class="MsoNormal" style="margin:2.4pt;background:white"><span style="font-size:9.0pt;color:#212529">The data in this note reflect different sub-national geographic levels in OECD countries:</span>
<ul style=“list-style-type:circle;font-size:9.0pt;color:#212529”>
<li><strong>Regions</strong>: are classified on two territorial levels reflecting the administrative organisation of countries: large regions (TL2) and small regions (TL3). Small regions are classified according to their access to metropolitan areas (Fadic et al. 2019).</li>
<li><strong>Functional urban areas</strong> consist of cities – defined as densely populated local units with at least 50 000 inhabitants – and adjacent local units connected to the city (commuting zones) in terms of commuting flows (Dijkstra, Poelman, and Veneri 2019). Metropolitan areas refer to functional urban areas above 250 000 inhabitants.</li></ul></p>

<p class="MsoNormal" style="margin:2.4pt;background:white"><span style="font-size:9.0pt;color:#212529">In addition, some indicators use the degree of urbanisation classification (OECD et al. 2021), which defines three types of areas:</span>
<ul style=“list-style-type:circle;font-size:9.0pt;color:#212529”>
<li><strong>Cities</strong> consist of contiguous grid cells that have a density of at least 1 500 inhabitants per km2 or are at least 50% built up, with a population of at least 50 000.</li>
<li><strong>Towns and semi-dense areas</strong> consist of contiguous grid cells with a density of at least 300 inhabitants per km2 and are at least 3% built up, with a total population of at least 5 000.</li>
<li><strong>Rural areas</strong> are cells that do not belong to a city or a town and semi-dense area. Most of these have a density below 300 inhabitants per km2.</li></ul></p>

<p class="MsoNormal" style="margin:2.4pt;background:white"><span style="font-size:9.0pt;color:#212529">Disclaimer: <a href="https://oecdcode.org/disclaimers/territories.html" target="_blank"><i><span style="color:#0d6efd">https://oecdcode.org/disclaimers/territories.html</span></i></a></span><u></u><u></u></p>
</td>
</tr>
</tbody>
</table>
