```{r isr_ctry}
ctry <- "ISR"
ctry2 <- "Israel"
```

# Israel

```{r isr_pkgs}
library(oecdcountryprofiles)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(oecdplot)
library(knitr)
library(readxl)
library(janitor)
library(stringr)
library(readr)
library(tintin)

load_oecd_fonts()

missing_reg <- c("ITG2F")
excl_countries <- c("LUX", "IRL")

clrs <- tintin_colours$the_black_island
clrs2 <- tintin_colours$the_blue_lotus
clrs3 <- tintin_colours$the_red_sea_sharks
clrs4 <- tintin_colours$the_calculus_affair
clrs5 <- tintin_colours$red_rackhams_treasure
```

<table style="margin-left:66.1pt;border-collapse:collapse" cellspacing="0" cellpadding="0" border="0">
<tbody>
<tr>
<td style="width:467.5pt;border:solid windowtext 1.0pt;padding:0cm 5.4pt 0cm 5.4pt" width="623" valign="top">
<p class="MsoNormal" style="background:#fff9e6">
<b><span style="font-size:9.0pt;color:#212529">OECD Regional Outlook</span></b><u></u><u></u></p>
<p class="MsoNormal" style="margin:2.4pt;background:white"><span style="font-size:9.0pt;color:#212529">The OECD
<a href="https://www.oecd-ilibrary.org/urban-rural-and-regional-development/oecd-regional-outlook_2dafc8cf-en" target="_blank"><i><span style="color:#0d6efd">Regional Outlook</span></i></a> reviews recent trends, policy developments, and prospects across OECD regions, including the underlying causes driving regional inequalities in performance and well-being. The report offers evidence,
 guidance and policy recommendations on how to improve competitiveness and productivity, promote inclusive growth, accelerate the net-zero transition and raise well-being standards through effective regional development policy and multi-level governance.</span><u></u><u></u></p>
</td>
</tr>
</tbody>
</table>

## Overview

```{r isr_tbl}
read_html_table(ctry)
```

## Recent policy developments

```{r isr_txt}
read_html_text(ctry)
```

<table border="1" style="width:75%;margin-left:auto;margin-right:auto;">
  <tr>
    <th style="background:#d4edff; padding:10px;font-size:9.0pt;" valign="center">
      Territorial definitions
    </th>
  </tr>
  <tr>
    <td>
      <p style="padding-left:10px;padding-right:10px;padding-top:10px;font-size:9.0pt;">
        The data in this note reflect different sub-national geographic levels in OECD countries. In particular, <b>regions</b> are classified on two territorial levels reflecting the administrative organisation of countries: large regions (TL2) and small regions (TL3).
      </p>
      <p style="padding-left:10px;padding-right:10px;font-size:9.0pt;">
        Small regions are classified according to their access to metropolitan areas (Fadic et al. 2019). The typology classifies small (TL3) regions into metropolitan and non-metropolitan regions according to the following criteria:
      </p>
      <ul style="list-style-type:circle;font-size:9.0pt;padding-right:10px">
        <li>
          <b>Metropolitan regions</b>, if more than half of the population live in a FUA. Metropolitan regions are further classified into: <b>metropolitan large</b>, if more than half of the population live in a (large) FUA of at least 1.5 million inhabitants; and <b>metropolitan midsize</b>, if more than half of the population live in a (midsize) FUA of at 250 000 to 1.5 million inhabitants.
        </li>
        <li>
          <b>Non-metropolitan regions</b>, if less than half of the population live in a midsize/large FUA. These regions are further classified according to their level of access to FUAs of different sizes: <b>near a midsize/large FUA</b> if more than half of the population live within a 60-minute drive from a midsize/large FUA (of more than 250 000 inhabitants) or if the TL3 region contains more than 80% of the area of a midsize/large FUA; <b>near a small FUA</b> if the region does not have access to a midsize/large FUA and at least half of its population have access to a small FUA (i.e. between 50 000 and 250 000 inhabitants) within a 60-minute drive, or contains 80% of the area of a small FUA; and <b>remote</b>, otherwise.
        </li>
      </ul>
      </p>
      <p style="padding-left:10px;padding-right:10px;font-size:9.0pt;">
        Disclaimer: <a href="https://oecdcode.org/disclaimers/territories.html" target="_blank"><i style="color:#0d6efd">https://oecdcode.org/disclaimers/territories.html</i></a>
      </p>
    </td>
  </tr>
</table>
